\chapter{Introduction}
\label{ch:Introduction}

\todo[inline]{FINAL TODOs:\\
- Search and Replace for CS and LD. Only the very first occurrence can be italics and capitalized, all other instances should be normal text\\
-- Citizen Science -> citizens science (no italics, just normal text)\\
-- Liquid Democracy -> liquid democracy \\
-- When using CS or LD make sure to wrap it in a \textbackslash{}tracknshrink\{CS/LD\}
}

In this project report we are linking \textit{Citizen Science} (\tracknshrink{CS}) and \textit{Liquid Democracy} (\tracknshrink{LD}). Both concepts share the same people-driven foundation, as both involve the participation of interested citizens and users likewise. Citizen science describes scientific research conducted by lay volunteers. In other words \textit{“Amateur Experts”}  \parencite{Gura2013}. The number of \tracknshrink{CS} projects and their uses in primary research have grown rapidly over the last years \parencite{Kosmala2016} and an increasing amount of today’s professional research relies on citizen scientist. Liquid democracy describes a compound of two fundamental types of democracy, that are (1) direct democracy and (2) representative or indirect democracy. That means, liquid democracy empowers citizens to either vote directly for a certain state of affairs or to  delegate their vote to another citizen. Beside both concepts share the same participatory nature, and despite being old ideas both of them shine in a digital environment. Interlinking these two conceptions is, however, a novel approach, and the primary objective of this project.

% This document is the project report of the  project group, a project in the \href{http://www.dh.uni-leipzig.de/wo/courses/summer-semester-20142015/citizen-science/}{\textit{Citizen Science}} course of the \href{http://www.dh.uni-leipzig.de/wo/}{department of Digital Humanities} at the University Leipzig in the summer semester of 2017.

\section{Motivation}
\label{sec:Motivation}

Participants of citizen science programs take an active role in a scientific research process. They usually work with other fellow citizen scientists towards the same goal and are often supervised---to some degree---by scientists, labs or institutions who initially started the study. Ultimately, citizen science projects entail multiple advantages for all stakeholders. For one, some research questions are simply impossible to answer without having an active mass helping to conduct, measure and interpret the data. Apart from that, this form of ‘distributed intelligence’ is a salient feature for \tracknshrink{CS}. On the other hand, citizen scientist gain additional knowledge in their field from other fellow peers. They strengthen their skills in communication, reporting, and scientific practices. There is also a motivational difference between a citizen scientist and a professional one. Since citizen scientists work voluntary in their field of interest, they don’t need be convinced, as they already have an affection for it. This may also yield in a different attitude towards their work and duties, and one could argue this ultimately leads to higher willingness to work.

Liquid democracy may also offer many opportunities for societies, whether for smaller communities, countries or for unions of states (e.g., the \tracknshrink{EU}). Since it does scale very well it has been successfully tested for smaller associations (e.g., Berlin’s Pirate Party). Since---simply put---liquid democracy combines direct and representative democracy, it offers whole new ways for the democratic participation process. This drastic shift would grant every citizens a natural right be heard, as everyone could use their voice for a certain state of affairs.

Since both concepts fully leverage their strengths in a digital environment, combining these concepts would further allow a liquid democracy system that is open for a wide field of research. Possible research fields would contain: psychology, sociology, demographics, and---apparently---political science, especially psephology (such as voter transition analysis). Not only researcher would benefit from such a system but the society as well, since it would change the entire process of decision-making and the way track political and historical events. 


\section{Objective}
\label{sec:Objective}

Although there are a few projects that implemented \tracknshrink{LD}---to various degrees---there is no project that would allow to access some of its data. Additionally, we focus on free and open-source software (\tracknshrink{FOSS}) solutions, since this would  

Despite a few organizations (ref) using them for participative processes, no platform allowing for processes that truly allow for Liquid Democracy exists (as we will show in \ref{sec:RelatedWork}, the tools we found are basically glorified voting tools). 

Furthermore, existing tools do not explicitly and readily make relevant scientific data transparent, and don't provide even the slightest assistance to citizens interested in understanding the processes taking place through them. 

We believe that it is time for a platform that truly allows for providing the structure for liquid democratic processes that are comprehensible for the centrepiece of a democratic system, the mature, participative Citizen.

\todo[inline]{Framework wie man Grundsätze einer Demokratie wahrn, reale Abbildung gesellschaftlicher Strukturen. Forschungsfragen stellen aus Daten und aus Community(!)}

As noted in the \hyperref[sec:Motivation]{motivation}, we believe in the importance of Citizen Science enabling liquid democracy within a participative system. Due to the transformation of processes of all kind into the digital world and more and more social context taking place in-silico, we believe this needs to take place through a participatory, decentralized and open online-platform.

The need for participation in societal processes becomes ever more apparent in light of the current developments of our political system. Ever more people feel disenfranchised, with little opportunity to make their voice heard on a political stage. Paradoxically, this development finds its expression in governments that allow for less participation and more centralized structures.

We believe that this not merely a technical problem, in the sense of lacking digital infrastructure; However we think that a societal movement empowering 'the citizen' can profit from a participatory, discursive platform.

As will be shown in \ref{ch:Conclusion}, such a platform does not exist. Our goal is thus to develop technological components, thoroughly grounded in the theory of both \hyperref[sec:Theory_CS]{Citizen Science} and \hyperref[sec:Liquid_Democracy]{Liquid Democracy} that provide a basis for interested citizens to develop such platforms.

These technological components need to enable the citizen to both participate in democratic processes in a way that respects its nature as mature political subject, as well as in enabling them to approach it as an object of research from a scientific perspective, even without being an expert in either the subject matter or research methods. 

It thus needs to provide the functionalities to exhibit the processes of a fully-functional, comprehensive liquid democracy process, as well as the ease of access and learning resources, as well as research infrastructure to perform Citizen Science that deserves its name.



without compromising the privacy and security of the participants.

\section{Structure}
\label{sec:structure}


After \hyperref[sec:Motivation]{motivating} the context of this research and \hyperref[sec:Objective]{formulating the research question}, the \hyperref[ch:Theory]{next section} will give discuss more thoroughly the two major societal contexts this project is situated in, namely \hyperref[sec:Theory_CS]{Citizen Science} and \hyperref[sec:Liquid_Democracy]{Liquid Democracy}. From the discussion of these theoretic aspects, we will derive a \hyperref[sec:Criteria]{range of criteria} a Citizen Science-enabled Liquid Democracy platform needs to fulfill.

Subsequently, we will explore existing approaches for Liquid Democracy platforms (or voting platforms as is often the case), presenting the current state of the art. We will evaluate in how far these platforms conform to the \hyperref[sec:Criteria]{criteria derived from the theory}. 

As will be shown, no existing approach is suited to fulfill the criteria. Thus, in \hyperref[ch:ProjectRequirements]{the following chapter} we will develop a platform based on the requirements sketched before. This will be done by \hyperref[sec:SoftwareRequirements]{compiling product requirements} the platform needs to fulfill in order to meet the projects requirements. Subsequently, we will explore options in the realms of \hyperref[sec:ArchitecturalDesign]{architectural design}, \hyperref[sec:ApplicationDesign]{application design} and provide protocols for (1) \hyperref[sec:PublicDeclarationOfIntentProtocol]{verifiable voting and delegation in a public space} and (2) \hyperref[sec:VoteTallyingProtocol]{vote tallying in the public space}.

Finally, we will \hyperref[ch:Conclusion]{conclude} by critically discussing the \hyperref[sec:DiscussionImplementation]{prototype implementation} in the context of the project, with a particular focus on the \hyperref[sec:Criteria]{criteria derived from the theory}, before we will conclude with a discussion of \hyperref[sec:FutureWork]{future work}.